---
title: "Network Status and Centrality"
description: |
  Who are the key actors in the football network? What makes them the key actors
author:
  - name: Isha Akshita Mahajan,Ankit Kumar
    affiliation: UMass Amherst
date: 2022-05-09
output:
  distill::distill_article:
    self_contained: false
---

## Assignment Description

Calculate closeness, betweenness and eigenvector centrality measures for your network data, and bonachic-power if possible. Compare these measures to basic degree centrality measures. Try to interpret the results. Are there any interesting patterns in the distribution of measures or correlations between them that provide insight into the measures?

You may also want to identify whether the same node(s) are more/less central using the different measures, and see if you can find any patterns in the results based on which nodes are high/low on each measures. Discuss (with any related evidence) whether or not the node(s) behavior is in line with or violates expectations based on the degree centrality measure, comparing across those measures using a data frame similar to that constructed in the syntax. Be sure to share your assignment with group members for comments.


## Eigenvector Centrality

The eigenvector centrality determines how important the neighborhood of the nodes is in the network. The higher the eigenvalue centrality, the more important the neighbors of that node are with respect to other nodes in the network. (Jackson 08) The eigenvalues also take into account the proximity of the node to many other important nodes in the network. 

AS Roma has the highest eigenvalue in the network. This can be an additional support to our previous speculation on how AS Roma might influential in the influx and outflow of football talent as they are well connected to the other influential nodes in our football transfer network. 


```{r include=FALSE}
library(igraph)
library(tidyverse)
library(corrr)
library(intergraph)
library(statnet)
library(network)
library(sna)
library(ggplot2)
library(kableExtra)

#load data from CSV
transfers <- read_csv("/Users/isha/Desktop/GitHub/transfernetworks.csv")
head(transfers)
#select relevant data for edgelist format
data <- transfers %>% 
  select(club_from, club_to, name,fee) %>% 
  filter(fee != "free transfer") %>% 
  filter(fee != "loan transfer") %>% 
  filter (fee != 0)
data$fee <- as.numeric(data$fee) 
data <- data %>% 
filter(fee != is.na(fee))
any(is.na(data$fee))
#convert data into matrix format
data.mat <- as.matrix(data)
#create igraph object from data
ig <- graph_from_data_frame(data.mat, directed = TRUE)
# add edge attribute weight i.e transfer fee
ig <- set_edge_attr(ig, "weight", value = na.omit(data$fee))
# add edge attribute season of transfer
#ig<-  set_edge_attr(ig, "season", value = na.omit(data$season))
# add node attribute i.e league the club belongs to
#set_vertex_attr(ig,"league" ,value = node_attr)
#delete edge attribute that was automatically being created in addition to weight
ig <- delete_edge_attr(ig, "fee")
#check summary of the igraph object
summary(ig)
#convert ig network into intergraph object to coerce with statnet
network <- intergraph::asNetwork(ig) 
#create degree column
transfer.nodes<-data.frame(name=V(ig)$name, degree=igraph::degree(ig))
#create indegree and outdegree column
transfer.nodes<-transfer.nodes %>%
           mutate(indegree=igraph::degree(ig, mode="in", loops=FALSE),
           outdegree=igraph::degree(ig, mode="out", loops=FALSE))
```


```{r warning=FALSE}
temp <- igraph::centr_eigen(ig,directed=T)
transfer.nodes$eigen <- temp$vector
#arrange descending and return top 5 nodes
arrange(transfer.nodes, desc(eigen)) %>%
  slice(1:5)
```

## Derived and Reflected Centrality

 We compute the derived and reflected Eigenvector centrality to better understand whether the influence is more observed in the neighborhood of the nodes or is it more observed on the strength of the node itself. 

```{r}
matrix <- as.matrix(as_adjacency_matrix(ig, attr = "weight"))
#square the adjacency matrix
matsq<-t(matrix) %*% matrix
#Calculate the proportion of reflected centrality.
transfer.nodes$rc<-diag(matsq)/rowSums(matsq)
#replace missing values with 0
transfer.nodes$rc<-ifelse(is.nan(transfer.nodes$rc),0,transfer.nodes$rc)
#Calculate received eigenvalue centrality
transfer.nodes$eigen.rc<-transfer.nodes$eigen*transfer.nodes$rc
#Calculate the proportion of derived centrality.
transfer.nodes$dc<-1-diag(matsq)/rowSums(matsq)
#replace missing values with 0
transfer.nodes$dc<-ifelse(is.nan(transfer.nodes$dc),1,transfer.nodes$dc)
#Calculate received eigenvalue centrality
transfer.nodes$eigen.dc<-transfer.nodes$eigen*transfer.nodes$dc
```

```{r}
filter(transfer.nodes,name%in%c("Chelsea FC","FC Barcelona"))
```

```{r}
##cen <- eigen_centrality(ig, directed = TRUE)
#cen$vector
```

```{r}
transfer.nodes%>%
  select(-name) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    geom_histogram() +
    facet_wrap(~key, scales = "free")
```

```{r}
#build on the dataframe
temp<-transfer.nodes %>% 
  select(degree,indegree,outdegree,eigen,eigen.rc,eigen.dc)%>%
  correlate() %>%
  rearrange()
```

```{r}
fashion(temp)
```

```{r}
rplot(temp)
```

## Betweeness Centrality

Betweeness centrality quantifies the number of times a node acts as a bridge along the shortest path between two other nodes.

```{r}
transfer.nodes$betweeness <- igraph::betweenness(ig, directed=TRUE)
transfer.nodes%>%
arrange(desc(betweeness))%>%
slice(1:5)
```

## Network Constraint 

```{r}
transfer.nodes$constraint<-constraint(ig)
```

## Authority 

Authorities are those actors which get many incoming links from hubs, presumably because of their high quality relevant information. (Klien 16) In this case Juventus, Chelsea, Barcelona,PSG, Manchester United and Athleteco have the highest authority scores. This means that these clubs are attracting talent from the hubs like Real Madrid, which we will see includes Eden Hazard's transfer in 2019

```{r}
transfer.nodes$authority <- authority_score(ig)$vector

a <- transfer.nodes %>% 
  arrange(desc(authority))
head(a)
```

## Hubs 

Many Ties to Peripheral Actors 

```{r}
transfer.nodes$hub_score <- hub_score(ig)$vector
head(transfer.nodes)
```

Consider Real Madrid a hub given that their hub_score is 1. We can confirm this by looking at higher eigenvector derived centrality and lower reflected centrality. 

## Bridges 

A few ties to central actors =

```{r}
bridges(ig)
bridges <- transfer.nodes %>% 
  filter(name == "Maccabi Haifa")
head(bridges)
```

We test the understanding whether these teams actually serve as bridges in the network. By looking at the derived and reflected centrality scores, we see that the the reflected centrality for Maccabi Haifa is low while the derived centrality is high, thereby making it a bridge. 


## Correlations for Centrality Measures 

```{r}
head(transfer.nodes)
cor<-transfer.nodes %>% 
  select(degree,eigen,betweeness, authority, hub_score)%>%
  correlate() %>%
  rearrange()
rplot(cor)
```

## Brokerage 

```{r}
tempo <-as.data.frame(brokerage(network, cl  = get.vertex.attribute(network,"league"))$raw.nli)
head(tempo)
transfer.nodes<-transfer.nodes %>%
mutate(broker.tot = tempo$t, # AS ROMA, AS Monaco, US Sassuolo, Genoa CFC, Atlanta BC
       broker.coord = tempo$w_I,
       broker.itin = tempo$w_O, # Aston Villa, Inter Milan, Borussia Dortmund, ACF florentina
       broker.rep = tempo$b_IO, # # Aston Villa, Inter Milan, Borussia Dortmund, ACF florentina
       broker.gate = tempo$b_OI,
       broker.lia = tempo$b_O)
transfer.nodes %>% 
  arrange(desc(broker.tot))
```

AS Roma plays a role of coordinator in the network which means that many ties go to and from AS Roma. This is in alignment with the degrees of the network as we saw earlier and the Eigenvector centrality which was also the highest for AS Roma in the network. 

Aston Villa is an iterant broker in the network. 





