---
title: "Exploring European Football Transfer Networks"
description: |
  Exploring a network of football teams and the transactions they made from 2018-2021. 
  An Edgelist maybe?
author:
  - name: Isha Akshita Mahajan
    affiliation: UMass Amherst
date: 2022-05-08
output:
  distill::distill_article:
    self_contained: false
---


## Load Required Packages 

```{r}
library(igraph)
library(tidyverse)
library(corrr)
library(intergraph)
library(statnet)
library(network)
library(sna)
library(ggplot2)
```

## Load the data 

I begin by converting the data into an edgelist format. I select the club from and club to columns which are our nodes in this case. The relationship that I'm interested in is where from and where to has the majority of transfers taken place across various places. After selecting the from and to columns, I select the players, the transfer fee and the season in which the transfer took place. I also remove loan and free transfer as they don't seem relevant at the moment. 

```{r}
#load data from CSV
transfers <- read_csv("/Users/isha/Desktop/GitHub/transfers2018-2021.csv")
#select relevant data for edgelist format
data <- transfers %>% 
  select(club_from, club_to, player,fee, season) %>% 
  filter(fee != "loan transfer") %>% 
  filter(fee != "free transfer") %>% 
  arrange(desc(season))
#convert fee to numeric character 
data$fee <- as.numeric(data$fee)
#convert season to numeric character 
data$season <- as.numeric(data$season)
#check summary for numeric conversion
summary(data)
```


```{r}
#create league names as node attributes
node1 <- transfers$league_from
node2 <- transfers$league_to
nodes <- c(node1, node2)
node_attr <- unique(nodes)
```

## Create Edgelist

The edgelist I created consists of 48 nodes and 7520 ties. the edge attributes are the player who is being transferred, and the season in which they were transferred. The weight of the edges is the fee for which they were transferred. The edgelist is directed because the players are transferring from one team to another. 

```{r}
#convert data into matrix format
data.mat <- as.matrix(data)
#create igraph object from data
ig <- graph_from_data_frame(data.mat, directed = TRUE)
# add edge attribute weight i.e transfer fee
ig <- set_edge_attr(ig, "weight", value = na.omit(data$fee))
# add edge attribute season of transfer
ig<-  set_edge_attr(ig, "season", value = na.omit(data$season))
# add node attribute i.e league the club belongs to
set_vertex_attr(ig,"league" ,value = node_attr)
#delete edge attribute that was automatically being created in addition to weight
delete_edge_attr(ig, "fee")
#check summary of the igraph object
summary(ig)
```

```{r}
#convert ig network into intergraph object to coerce with statnet
network <- intergraph::asNetwork(ig) 
```

```{r}
#plot intergraph object
plot(network)
```

```{r}
#plot igraph object 
plot(ig)
```

## Exploring Network Structures 

```{r}
#count the number of nodes
vcount(ig)
# count the number of edges
ecount(ig)
```

The network consists of 48 nodes and 7520 edges. 

```{r}
# look at the dyad census
dyad_census(ig)
```

**There are 3 mutual, 81 asymmetrical and 1044 dyads. Elaborate Explanation**

```{r}
triad_census(ig)
```

**Add Explanation**

```{r}
is_directed(ig)
is_weighted(ig)
is_bipartite(ig)
```

The results suggest that the network is directed, i.e players are transferring from one club to another. 

The network is weighted. The weight of the transfer is the fee for which the transfer was made. 

The network is not bipartite which mean that they transfers are not in sets- there is a flow of players. 


## Transitivity

```{r}
#get global clustering cofficient: igraph
transitivity(ig, type="global")
#get average local clustering coefficient: igraph
transitivity(ig, type="average")
```


The global transitivity of the graph is 0.1598063 which is the ratio of triangles connected to triangles

the average transitivity is the transitivity of the local triad clusters, i.e. the ratio of local triangles to all connected triangles. 

## Path Lengths 

The average path length in the weighted network is 164480670. **Add Interpretation**

```{r}
average.path.length(ig)
```

The distances function enables us to look at the distance travelled between two nodes. Let's explore some distances between football clubs of various leagues.

```{r}
distances(ig,"Liverpool FC", "FC Porto")
```

The distance between Liverpool FC and FC Porto is approximitely 5 nodes. 

```{r}
distances(ig,"Chelsea FC", "Real Madrid")
```

The distance between Chelsea FC and Real Madrid is 3.8 nodes. 


```{r}
distances(ig, "Chelsea FC", "FC Barcelona")
```

The distance between Chelsea and FC Barcelona is 1.6 nodes.

## Component Structure 

```{r}
names(igraph::components(ig))
igraph::components(ig)$no
igraph::components(ig)$membership
igraph::components(ig)$csize
```

There are no subcomponents in the graph. 


