---
title: "Network Degree"
description: |
  Network Degree of Football Networks
author:
  - name: Isha Akshita Mahajan
    affiliation: UMass Amherst
date: 2022-05-08
output:
  distill::distill_article:
    self_contained: false
---

## Density 

```{r}
graph.density(ig, loops = FALSE)
```

The network density of the football transfer network is 0.005 which indicates that the 0.5% of all possible ties in the network are made. In this case we would leave the loops value as false because it is not possible for a football team to transfer a player back to their own team. 

## Degree

The degree measures the node centrality or a popularity of a node. It gives a count of the relationships that a node is involved in or the number of edges it has. Given that the goal of these blogs to keep all these important network properties in one place, I begin by creating a dataframe that I'll keep on building as I create these blogs. I call this data frame `transfer.nodes`. 

```{r}
#create degree column
transfer.nodes<-data.frame(name=V(ig)$name, degree=igraph::degree(ig))
#create indegree and outdegree column
transfer.nodes<-transfer.nodes %>%
           mutate(indegree=igraph::degree(ig, mode="in", loops=FALSE),
           outdegree=igraph::degree(ig, mode="out", loops=FALSE))
```


```{r}
# clubs that have the highest relationships in the network
degree <- transfer.nodes %>% 
  arrange(desc(degree)) 
kable(head(degree))
# Clubs that recieved the most transfers.
indegree <- transfer.nodes %>% 
  arrange(desc(indegree)) 
kable(head(indegree))
# Clubs that transferred the most.
outdegree <- transfer.nodes %>% 
arrange(desc(outdegree)) 
kable(head(outdegree))
```

AS ROMA has the highest degree in our football transfer network. Both the in degree and the out degrees hold a value of 39. This means that 39 transfer of players have left the club and 39 players have come in to the club.This makes me wonder if AS ROMA is a club which might be serving like a starting point for various players to move towards the major football leagues. But there might be better ways to explore that later in the analysis. 

Genoa FC has the indegree of 46 which means they've had 46 transfers to their club. 

Juventus FC has the highest ourdegree which means they've transferred 41 players out of the club. 


```{r}
#V(ig)$name
roma_movement <- incident(ig, 66, mode = c("all"))
```

```{r}
table(head_of(ig, E(ig)))
```

```{r}
# make a basic plot
plot(ig, 
     vertex.label.color = "black", 
     edge.color = 'gray77',
     vertex.size = 6,
     edge.arrow.size = 0.1,
     layout = layout_nicely(ig))
```


## Degree Distributions

```{r}
#create a histogram of Transfer Indegree
hist(transfer.nodes$indegree, main="Football Transfers: In-degree Distribution", xlab="Players Recieved")
```

```{r}
#create a histogram of Transfer Indegree
hist(transfer.nodes$outdegree, main="Football Transfers: Out-degree Distribution", xlab="Players Transferred")
```

## Network Degree Centralization

**Add Interpretation, check if loops are supposed to be true or false** 

```{r}
#get network centralization score: igraph
centr_degree(ig, loops = FALSE, mode="in")$centralization
centr_degree(ig, loops = FALSE, mode="out")$centralization
```


#Degree, Eigen Vector, Hub, Authority, closeness, betweeness - Centrality
