---
title: "Network Degree"
description: |
  Network Degree of Football Networks
author:
  - name: Isha Mahajan, Ankit Kumar
    affiliation: UMass Amherst
date: 2022-05-08
output:
  distill::distill_article:
    self_contained: false
---


## Assignment Description

Briefly describe the dataset you are using: identify initial network format, describe and identify the nodes (including how many nodes are in the dataset), what constitutes a tie or edge (including how many ties, whether ties are directed/undirected and weighted/binary, and how to interpret the value of the tie if any), whether or not there are edge attributes that might be used to subset data or stack multiple networks (e.g., tie type, year, etc). Not every feature of the network needs to be described, but description should orient reader to the network data and provide any necessary context for the results provided.

Provide at least two or three noteworthy results, including the relevant statistics and interpretation. For example, explaining which node(s) are most central and which are least central. Discuss (with any related evidence) whether or not the node(s) behavior is in line with or violates expectations based on the degree centrality measure. What do you make of network density and centralization measures?


```{r include=FALSE}
library(igraph)
library(tidyverse)
library(corrr)
library(intergraph)
library(statnet)
library(network)
library(sna)
library(ggplot2)
library(kableExtra)

#load data from CSV
transfers <- read_csv("/Users/isha/Desktop/GitHub/transfernetworks.csv")
head(transfers)
#select relevant data for edgelist format
data <- transfers %>% 
  select(club_from, club_to, name,fee) %>% 
  filter(fee != "free transfer") %>% 
  filter(fee != "loan transfer") %>% 
  filter (fee != 0)
data$fee <- as.numeric(data$fee) 
data <- data %>% 
filter(fee != is.na(fee))
any(is.na(data$fee))
#convert data into matrix format
data.mat <- as.matrix(data)
#create igraph object from data
ig <- graph_from_data_frame(data.mat, directed = TRUE)
# add edge attribute weight i.e transfer fee
ig <- set_edge_attr(ig, "weight", value = na.omit(data$fee))
# add edge attribute season of transfer
#ig<-  set_edge_attr(ig, "season", value = na.omit(data$season))
# add node attribute i.e league the club belongs to
#set_vertex_attr(ig,"league" ,value = node_attr)
#delete edge attribute that was automatically being created in addition to weight
ig <- delete_edge_attr(ig, "fee")
#check summary of the igraph object
summary(ig)
#convert ig network into intergraph object to coerce with statnet
network <- intergraph::asNetwork(ig) 
```


## Density 

```{r collapse=TRUE}
graph.density(ig, loops = FALSE)
```

The network density of the football transfer network is 0.005 which indicates that the 0.5% of all possible ties in the network are made. In this case we would leave the loops value as false because it is not possible for a football team to transfer a player back to their own team. 

## Degree

The degree measures the node centrality or a popularity of a node. It gives a count of the relationships that a node is involved in or the number of edges it has. Given that the goal of these blogs to keep all these important network properties in one place, I begin by creating a dataframe that I'll keep on building as I create these blogs. I call this data frame `transfer.nodes`. 

```{r collapse=TRUE}
#create degree column
transfer.nodes<-data.frame(name=V(ig)$name, degree=igraph::degree(ig))
#create indegree and outdegree column
transfer.nodes<-transfer.nodes %>%
           mutate(indegree=igraph::degree(ig, mode="in", loops=FALSE),
           outdegree=igraph::degree(ig, mode="out", loops=FALSE))
```


```{r collapse=TRUE}
# clubs that have the highest relationships in the network
degree <- transfer.nodes %>% 
  arrange(desc(degree)) 
kable(head(degree))
# Clubs that recieved the most transfers.
indegree <- transfer.nodes %>% 
  arrange(desc(indegree)) 
kable(head(indegree))
# Clubs that transferred the most.
outdegree <- transfer.nodes %>% 
arrange(desc(outdegree)) 
kable(head(outdegree))
```

AS ROMA has the highest degree in our football transfer network. Both the in degree and the out degrees hold a value of 39. This means that 39 transfer of players have left the club and 39 players have come in to the club.This makes me wonder if AS ROMA is a club which might be serving like a starting point for various players to move towards the major football leagues. But there might be better ways to explore that later in the analysis. 

Genoa FC has the indegree of 46 which means they've had 46 transfers to their club. 

Juventus FC has the highest ourdegree which means they've transferred 41 players out of the club. 

```{r collapse=TRUE}
#V(ig)$name
roma_movement <- incident(ig, 66, mode = c("all"))
```

```{r collapse=TRUE}
#table(head_of(ig, E(ig)))
```

```{r collapse=TRUE}
# make a basic plot
plot(ig, 
     vertex.label.color = "black", 
     edge.color = 'gray77',
     vertex.size = 6,
     edge.arrow.size = 0.1,
     layout = layout_nicely(ig))
```


## Degree Distributions

```{r collapse=TRUE}
#create a histogram of Transfer Indegree
hist(transfer.nodes$indegree, main="Football Transfers: In-degree Distribution", xlab="Players Recieved")
```

```{r collapse=TRUE}
#create a histogram of Transfer Indegree
hist(transfer.nodes$outdegree, main="Football Transfers: Out-degree Distribution", xlab="Players Transferred")
```

## Network Degree Centralization

**Add Interpretation, check if loops are supposed to be true or false** 

```{r collapse=TRUE}
#get network centralization score: igraph
centr_degree(ig, loops = FALSE, mode="in")$centralization
centr_degree(ig, loops = FALSE, mode="out")$centralization
```

